The Role of Variability in Learning Generalization: A Computational Modeling Approach

### 2024-10-12 Sections/full.qmd
few new references, and a re-render with the intro figure.

::: {.content-visible when-format="pdf"}
{{< include /Sections/pre_sections/toc_pdf.qmd >}}
:::

@@

One of the earliest and still often cited investigations of Schmidt's benefits of variability hypothesis was the work of @kerrSpecificVariedPractice1978. Two groups of children, aged 8 and 12, were assigned to either constant or varied training of a bean bag throwing task. The constant group practiced throwing a bean-bag at a small target placed 3 feet in front of them, and the varied group practiced throwing from a distance of both 2 feet and 4 feet. Participants were blindfolded and unable to see the target while making each throw but would receive feedback by looking at where the beanbag had landed in between each training trial. 12 weeks later, all of the children were given a final test from a distance of 3 feet which was novel for the varied participants and repeated for the constant participants. Participants were also blindfolded for testing and did not receive trial by trial feedback in this stage. In both age groups, participants performed significantly better in the varied condition than the constant condition, though the effect was larger for the younger, 8-year-old children. This result provides particularly strong evidence for the benefits of varied practice, as the varied group outperformed the constant group even when tested at the "home-turf" distance that the constant group had exclusively practiced. A similar pattern of results was observed in another study wherein varied participants trained with tennis, squash, badminton, and short-tennis rackets were compared against constant subjects trained with only a tennis racket [@greenPracticeVariabilityTransfer1995a]. One of the testing conditions had subjects repeat the use of the tennis racket, which had been used on all 128 training trials for the constant group, and only 32 training trials for the varied group. Nevertheless, the varied group outperformed the constant group when using the tennis racket at testing, and also performed better in conditions with several novel racket lengths. However, as is the case with many of the patterns commonly observed in the "benefits of variability" literature, the pattern wherein the varied group outperforms the constant group even from the constants group's home turf has not been consistently replicated. One recent study attempted a near replication of the Kerr & Booth study [@willeyLongtermMotorLearning2018], having subjects throw beanbags at a target, with the varied group training from positions (5 and 9 feet) on either side of the constant group (7 feet). This study did not find a varied advantage from the constant training position, though the varied group did perform better at distances novel to both groups. However, this study diverged from the original in that the participants were adults; and the amount of training was much greater (20 sessions with 60 practice trials each, spread out over 5-7 weeks).

@@

pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
  brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
  stringr, here,conflicted, patchwork, knitr, cowplot, RColorBrewer,data.table,gt,
  ggh4x,htmltools,future,furrr,ez)
#options(brms.backend="cmdstanr",mc.cores=4)
#options(digits=2, scipen=999, dplyr.summarise.inform=FALSE, kableExtra.auto_format=FALSE)
walk(c("brms","dplyr","bayestestR","here"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model", "prep_model_data","org_functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))
#conflicts_prefer(tidybayes::rstudent_t)

>>>>

::: {.content-visible when-format="pdf"}
<!-- {{< include /Sections/pre_sections/toc_pdf.qmd >}} -->
:::

@@

One of the earliest and still often cited investigations of Schmidt's benefits of variability hypothesis was the work of @kerrSpecificVariedPractice1978. Two groups of children, aged 8 and 12, were assigned to either constant or varied training of a bean bag throwing task. The constant group practiced throwing a bean-bag at a small target placed 3 feet in front of them, and the varied group practiced throwing from a distance of both 2 feet and 4 feet (see @fig-ex-design1). Participants were blindfolded and unable to see the target while making each throw but would receive feedback by looking at where the beanbag had landed in between each training trial. 12 weeks later, all of the children were given a final test from a distance of 3 feet which was novel for the varied participants and repeated for the constant participants. Participants were also blindfolded for testing and did not receive trial by trial feedback in this stage. In both age groups, participants performed significantly better in the varied condition than the constant condition, though the effect was larger for the younger, 8-year-old children. This result provides particularly strong evidence for the benefits of varied practice, as the varied group outperformed the constant group even when tested at the "home-turf" distance that the constant group had exclusively practiced. A similar pattern of results was observed in another study wherein varied participants trained with tennis, squash, badminton, and short-tennis rackets were compared against constant subjects trained with only a tennis racket [@greenPracticeVariabilityTransfer1995a]. One of the testing conditions had subjects repeat the use of the tennis racket, which had been used on all 128 training trials for the constant group, and only 32 training trials for the varied group. Nevertheless, the varied group outperformed the constant group when using the tennis racket at testing, and also performed better in conditions with several novel racket lengths. However, as is the case with many of the patterns commonly observed in the "benefits of variability" literature, the pattern wherein the varied group outperforms the constant group even from the constants group's home turf has not been consistently replicated. One recent study attempted a near replication of the  Kerr & Booth study [@willeyLongtermMotorLearning2018], having subjects throw beanbags at a target, with the varied group training from positions (5 and 9 feet) on either side of the constant group (7 feet). This study did not find a varied advantage from the constant training position, though the varied group did perform better at distances novel to both groups. However, this study diverged from the original in that the participants were adults; and the amount of training was much greater (20 sessions with 60 practice trials each, spread out over 5-7 weeks).

```{r}
#| label: fig-ex-design1
#| fig-width: 10
#| fig-height: 6
#| fig-cap: "A schematic representation of the @kerrSpecificVariedPractice1978 study design. The varied group trained from two distances (2 and 4 feet), while the constant group trained from a single distance (3 feet). Both groups were tested from a distance of 3 feet. The varied group outperformed the constant group at testing, despite the constant group having exclusively practiced from the testing distance."
#| cache: false

pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
  brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
  stringr, here,conflicted, patchwork, knitr, cowplot, RColorBrewer,data.table,gt,
  ggh4x,htmltools,future,furrr,ez)

create_data <- function(varied_train, constant_train, both_test) {
  varied_labels <- "Varied-Train"
  constant_labels <- "Constant-Train"
  test_labels <- "Testing"

  data <- data.frame(
    x = c(varied_train, constant_train, both_test),
    y = c(rep(0.55, length(varied_train)), rep(0.75, length(constant_train)), rep(0.4, length(both_test))),
    label = c(rep(varied_labels, length(varied_train)),
              rep(constant_labels, length(constant_train)),
              rep(test_labels, length(both_test))),
    color = c(rep("#00A08A", length(varied_train)), rep("#FF0000", length(constant_train)), rep("#ECCBAE", length(both_test)))
  )

  data
}

create_segments <- function(data) {
  min_x <- min(data$x)
  max_x <- max(data$x)
  segments <- data.frame(
    x = c(min_x - 0.5, min_x - 0.5, min_x - 0.5),
    xend = c(max_x + 0.5, max_x + 0.5, max_x + 0.5),
    y = c(0.55, 0.75, 0.4),
    yend = c(0.55, 0.75, 0.4)
  )
  segments
}

create_plot <- function(varied_train, constant_train, both_test) {
  data <- create_data(varied_train, constant_train, both_test)
  segments <- create_segments(data)

  # Variables for Training segment position
  train_x <- min(data$x) - .5
  train_y_start <- 0.5
  train_y_end <- 0.8
  train_text_y <- 0.75

  # Variables for Testing segment position
  test_x <- min(data$x) - .5
  test_y_start <- 0.35
  test_y_end <- 0.45
  test_text_y <- 0.48

  ggplot() +
    geom_segment(data = segments, aes(x = x, xend = xend, y = y, yend = yend),
                 linewidth = 1.2, color = "gray40") +
    geom_point(data = data, aes(x = x, y = y, color = color), size = 12) +
    geom_text(data = data, aes(x = x, y = y, label = label),
              color = "black", size = 4.5, vjust = -1.8, hjust = 0.5, fontface = "italic") +
    geom_segment(aes(x = train_x - 0.3, xend = train_x - 0.3, y = train_y_start, yend = train_y_end), color = "black", linewidth = 1.2) +
    geom_segment(aes(x = train_x - 0.5, xend = train_x, y = train_y_end, yend = train_y_end), color = "black", linewidth = 1.2) +
    geom_segment(aes(x = train_x - 0.5, xend = train_x, y = train_y_start, yend = train_y_start), color = "black", linewidth = 1.2) +
    geom_text(aes(x = train_x - .6, y = train_text_y, label = "Training"), angle = 90, vjust = 0.5, hjust = 1.5, size = 4.5, fontface = "bold") +
    geom_segment(aes(x = test_x - 0.3, xend = test_x - 0.3, y = test_y_start, yend = test_y_end), color = "black", linewidth = 1.2) +
    geom_segment(aes(x = test_x - 0.5, xend = test_x, y = test_y_end, yend = test_y_end), color = "black", linewidth = 1.2) +
    geom_segment(aes(x = test_x - 0.5, xend = test_x, y = test_y_start, yend = test_y_start), color = "black", linewidth = 1.2) +
    geom_text(aes(x = test_x - .6, y = test_text_y, label = "Testing"), angle = 90, vjust = 0.5, hjust = 1.5, size = 4.5, fontface = "bold") +
    scale_color_manual(values = c("#00A08A","#ECCBAE", "#FF0000")) +
    xlim(min(data$x) - 2, max(data$x) + 1) +
    ylim(0.3, 1) +
    theme_void() +
    theme(legend.position = "none",
          plot.margin = margin(10, 20, 10, 20),
          plot.background = element_rect(fill = "white", color = NA),
          panel.background = element_rect(fill = "white", color = NA))
}

kerr_booth_1978 <- create_plot(varied_train = c(2, 4), constant_train = c(3), both_test = c(3))
kerr_booth_1978
```

@@

# pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
#   brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
#   stringr, here,conflicted, patchwork, knitr, cowplot, RColorBrewer,data.table,gt,
#   ggh4x,htmltools,future,furrr,ez)
#options(brms.backend="cmdstanr",mc.cores=4)
#options(digits=2, scipen=999, dplyr.summarise.inform=FALSE, kableExtra.auto_format=FALSE)
walk(c("brms","dplyr","bayestestR","here"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model", "prep_model_data","org_functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))
#conflicts_prefer(tidybayes::rstudent_t)
